Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2016-10-26 paper.tex
Expand elastic scattering section to use J(eps) and H^a(eps)

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\todo{code these up}
We may take angular moments of the distribution function:
\begin{align}
  \label{eqn:J}
  J(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  \label{eqn:Ha}
  H^\mu(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
  \label{eqn:Sab}
  S^{\mu\gamma}(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \ell^\gamma,
\end{align}
representing specific energy density, specific energy flux, and
specific momentum flux, respectively.
Here ``specific'' refers to the quantity being integrable over neutrino energy.
\todo{relate moments to other frames?}
Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
momentum space while holding $\varepsilon$ constant:
$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
\todo{consistent with Eqn.~\ref{eqn:def_direction}?}

@@

\begin{table}%[h]
  \caption{
    We analyze neutrino interaction processes in terms of these categories.
    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
    $N$ represents a nucleon $n$ or $p$,
    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
    $\gamma$ represents a high-energy photon.
    A prime indicates a change in that particle's energy.
  }
  \label{tab:neutrino_processes}
  \begin{tabularx}{\columnwidth}{X X}
    \hline \hline
    absorption/emission
    & $\nu_e + n \leftrightarrow e^- + p$                          \\
    & $\bar{\nu}_e + p \leftrightarrow e^+ + n$                    \\
    & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
    \hline
    elastic scattering
    & $\nu + N \leftrightarrow \nu + N$                            \\
    & $\nu + {}^AZ \leftrightarrow \nu + {}^AZ$                    \\
    \hline
    inelastic scattering
    & $\nu + e^- \leftrightarrow \nu' + e^{-'}$                    \\
    & $\nu + e^+ \leftrightarrow \nu' + e^{+'}$                    \\
    \hline
    thermal pair processes
    & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
    & $\nu + \bar{\nu} + N + N \leftrightarrow N' + N'$            \\
    & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
    \hline \hline
  \end{tabularx}
\end{table}

@@

\subsection{Elastic scattering}
\label{ssec:sources_se}
Neutrino scattering on particles of mass much greater than $\varepsilon$
is essentially iso-energetic. Following 
\cite[Eqn.~A8]{brue1985-core_collapse} or
\cite[Eqn.~4.20]{shib2011-truncated_moment} the collision term for the
$i$-th process takes the form
\begin{equation}
  \label{eqn:integral_se}
  C_{{\rm SE},i}
  = \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega')
  \big(f(\ell'_\beta) - f(\ell_\alpha)\big)
\end{equation}
where $R(\omega')$ is the scattering kernel from direction $\ell'_\beta$
to direction $\ell_\alpha$ having dimension $[R]={\rm energy}^{-1}$,
and the cosine of the scattering angle is
$\omega' \equiv 1-\ell'_\alpha \ell^\alpha$.
We write this in the form
\begin{equation}
  \label{eqn:sources_se}
  C_{{\rm SE},i}
  = \mathscr{K}_{{\rm SE},i} \Phi_{{\rm SE},i} - \mathscr{K}_{{\rm SE},i} f,
\end{equation}
with
\begin{align}
  \label{eqn:se_factor}
  \Phi_{{\rm SE},i} &\equiv
  \frac{1}{\mathscr{K}_{{\rm SE},i}}
  \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega')f(\ell'_\beta) \\
  \label{eqn:se_opacity}
  \mathscr{K}_{{\rm SE},i} &\equiv
  \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega').
\end{align}

It is customary to approximate the scattering kernel to linear order in
$\omega$ (as in~\cite[Eqn.~4.21]{shib2011-truncated_moment}):
\begin{equation}
  \label{eqn:se_kernel_expansion}
  R(\varepsilon,\omega) \approx R_0(\varepsilon) + \omega R_1(\varepsilon),
\end{equation}
which allows us to write Eqns.~\ref{eqn:se_factor} and \ref{eqn:se_opacity} as
\begin{align}
  \label{eqn:se_factor_expansion}
  \Phi_{{\rm SE},i} &=  \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
  J(\varepsilon) -
  \frac{1}{\mathscr{K}_{{\rm SE},i}} \ell^\alpha H_\alpha(\varepsilon)\\
  \label{eqn:se_opacity_expansion}
  \mathscr{K}_{{\rm SE},i} &= 4\pi \frac{\varepsilon^3}{(2\pi)^3}
  \big(R_0(\varepsilon)+R_1(\varepsilon)\big).
\end{align}

>>>>

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\todo{code these up}
We may take angular moments of the distribution function:
\begin{align}
  \label{eqn:J}
  J(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  \label{eqn:Ha}
  H^\mu(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \\
  \label{eqn:Sab}
  S^{\mu\gamma}(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \ell'^\gamma,
\end{align}
representing specific energy density, specific energy flux, and
specific momentum flux, respectively.
Here ``specific'' refers to the quantity being integrable over neutrino energy.
\todo{relate moments to other frames?}
Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
momentum space while holding $\varepsilon$ constant:
$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
\todo{consistent with Eqn.~\ref{eqn:def_direction}?}

@@

\begin{table}%[h]
  \caption{
    We analyze neutrino interaction processes in terms of these categories.
    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
    $N$ represents a nucleon $n$ or $p$,
    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
    $\gamma$ represents a high-energy photon.
    A prime indicates a change in that particle's energy.
  }
  \label{tab:neutrino_processes}
  \begin{tabularx}{\columnwidth}{X X}
    \hline \hline
    absorption/emission
    & $\nu_e + n \leftrightarrow e^- + p$                          \\
    & $\bar{\nu}_e + p \leftrightarrow e^+ + n$                    \\
    & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
    \hline
    elastic scattering
    & $\nu + N \leftrightarrow \nu + N$                            \\
    & $\nu + {}^AZ \leftrightarrow \nu + {}^AZ$                    \\
    \hline
    inelastic scattering
    & $\nu + e^- \leftrightarrow \nu' + e^{-'}$                    \\
    & $\nu + e^+ \leftrightarrow \nu' + e^{+'}$                    \\
    \hline
    thermal pair processes
    & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
    & $\nu + \bar{\nu} + N + N \leftrightarrow N' + N'$            \\
    & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
    % & $\nu_\alpha + \bar{\nu}_\alpha \leftrightarrow \nu_\beta + \bar{\nu}_\beta $ \\
    \hline \hline
  \end{tabularx}
\end{table}

@@

\subsection{Elastic scattering}
\label{ssec:sources_se}
Neutrino scattering on particles of mass much greater than $\varepsilon$
is essentially iso-energetic. Following 
\cite[Eqn.~A8]{brue1985-core_collapse} or
\cite[Eqn.~4.20]{shib2011-truncated_moment} the collision term for the
$i$-th process takes the form
\begin{equation}
  \label{eqn:integral_se}
  C_{{\rm SE},i}
  = \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega')
  \big(f(\ell'_\beta) - f(\ell_\alpha)\big)
\end{equation}
where $R(\omega')$ is the scattering kernel from direction $\ell'_\beta$
to direction $\ell_\alpha$ having dimension $[R]={\rm energy}^{-1}$,
and the cosine of the scattering angle is
$\omega' \equiv \ell'_\alpha \ell^\alpha$.
We write this in the form
\begin{equation}
  \label{eqn:sources_se}
  C_{{\rm SE},i}
  = \mathscr{K}_{{\rm SE},i} \Phi_{{\rm SE},i} - \mathscr{K}_{{\rm SE},i} f,
\end{equation}
with
\begin{align}
  \label{eqn:se_factor}
  \Phi_{{\rm SE},i} &\equiv
  \frac{1}{\mathscr{K}_{{\rm SE},i}}
  \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega')f(\ell'_\beta) \\
  \label{eqn:se_opacity}
  \mathscr{K}_{{\rm SE},i} &\equiv
  \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega').
\end{align}

It is customary to approximate the scattering kernel to linear order in
$\omega$ (as in~\cite[Eqn.~4.21]{shib2011-truncated_moment}):
\begin{equation}
  \label{eqn:se_kernel_expansion}
  R_{{\rm SE},i}(\varepsilon,\omega) \approx
  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon),
\end{equation}
which allows us to write Eqns.~\ref{eqn:se_factor} and \ref{eqn:se_opacity} as
\begin{align}
  \label{eqn:se_factor_expansion}
  \Phi_{{\rm SE},i}(\varepsilon) &= \frac{1}{\mathscr{K}_{{\rm SE},i}}
  \big(R_{{\rm SE},i}^0(\varepsilon) J(\varepsilon) +
   R_{{\rm SE},i}^1(\varepsilon) \ell_\alpha H^\alpha(\varepsilon) \big) \\
  \label{eqn:se_opacity_expansion}
  \mathscr{K}_{{\rm SE},i}(\varepsilon) &=
  4\pi \frac{\varepsilon^3}{(2\pi)^3} R_{{\rm SE},i}^0(\varepsilon),
\end{align}
where we have made the energy-dependence explicit.

However, we don't know the energy dependence of $J(\varepsilon)$ and
$H^\alpha(\varepsilon)$,
since the the truncated moment simulations usually evolve only the integrated
moments, $J$ and $H^\alpha$.
